---
title: "Processing and Visualizing Spatial Data in R"
author: "Chengjun Hou"
date: "March 24, 2016"
output:
  html_document:
    theme: cosmo
---


---------------

This document covers the following items for working with spatial data in R:
  
- Get machine and data ready
- Data structure of spatial polygons introduced by **sp**
- Function `over()` for polygon overlays searching

---------------



## Get Machine and Data Ready 

First, install R of course. Even better, download Rstudio <http://www.rstudio.com/>, 
an incredibly efficient and easy to use interface for working with R.  

R does not support working with spatial data straight out of the box 
so there are a couple of packages that need to be downloaded to get R working with spatial data. 
The two packages required are **sp** and **maptools**, and **leaflet** for cool visualization. 
Use the following scripts to get the packages installed and loaded:
```{r, eval=FALSE}
install.packages(c("sp","maptools","leaflet"))
```
```{r, message=FALSE, warning=FALSE}
library(sp)
library(maptools)
library(leaflet)
```

Now it's time to load the data. Setting up a working directory where the raw data sits is always the first step.
Here we have two spatial data files that needed to be loaded into the memory from two different locations, 
so function `setwd()` is called twice.
```{r}
setwd("D:/Workspace/Tutorial/geo_data")
# load rain fall data of 20160101
RainShapeFiles <- readShapePoly("VClimateRainfall20160101_CONUS.shp")
# load US county data
CountyShapeFiles <- readShapePoly("cb_2014_us_county_500k.shp")
CountyShapeFiles$NAME <- as.character(CountyShapeFiles$NAME)
```



## Data Structure of Spatial Polygons

`RainShapeFiles` and `CountyShapeFiles` are *SpatialPolygonsDataFrame* objects. 
You can interpret this as the combination of polygon shape files and a data frame. 
Each polygon shape file contains the boundary coordinates, 
and the data frame contains other attributes attached to each polygon. 
`RainShapeFiles` has 926 polygons and 9 attributes attached to each polygon.
```{r}
# dimension check
dim(RainShapeFiles)
# attribute names
names(RainShapeFiles)
```

As a demo, here we see a region in the state of NY where it is raining.
```{r}
# selecting row 736 to 742 of RainShapeFiles
NYRainShapeFiles <- RainShapeFiles[736:742,]
# selecting counties from NY state
NYCountyShapeFiles <- CountyShapeFiles[CountyShapeFiles$STATEFP==36, ]
```

Then we can have a quick visualization of this region in the state of NY.
```{r}
V3 = leaflet() %>% 
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(data = NYCountyShapeFiles, popup=NYCountyShapeFiles$NAME,
              fillColor = "green", fillOpacity = 0.2,
              stroke = T, color="Black", weight=2, opacity=0.2) %>%
  addPolygons(data = NYRainShapeFiles,
              fillColor = "red", fillOpacity = 0.6,
              stroke = T, color="Black", weight=1, opacity=0.1)
V3
```


## Overlay Searching
Numerical "map overlay" combines spatial features from one map layer 
with the attribute (numerical) properties of another. 
This vignette explains the R method `over()`, which provides a consistent way to retrieve indices or attributes
from a given spatial object (map layer) at the locations of another spatial object.

Given two geometries, A and B, command `over(A,B)` retrieves the geometry (location) 
indices of B at the locations of A. In particular, 
an integer vector of length 'length(A)' is returned, 
with `NA` values for locations in A not matching with locations in B (e.g. those points outside a set of polygons).
With this function, we can have a list of counties that interact with all raining polygons.
```{r}
OverlayIndex = !is.na(over(NYCountyShapeFiles, NYRainShapeFiles)[,1])
NYCountyShapeFiles[["NAME"]][OverlayIndex]
```

Finally we can have a more detailed list of which counties are interacted with each polygon. 
The data frame shown in between `[[1]]` and `[[2]]` contains attributes of the counties that interact 
with the first raining polygon. And since the 4th polygon is in Lake Erie, 
there is no interaction with any county.
```{r}
over(NYRainShapeFiles, NYCountyShapeFiles, returnList = TRUE)
```



---------------