---
title: "HTML Slides Demo"
author: "Chengjun Hou"
date: "February 9, 2016"
output:
  ioslides_presentation:
    logo: HT.png
    widescreen: false
    smaller: false
    transition: faster
---



# Interactive Map

## Leaflet Demo

This is an R leaflet library example.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(maps)
library(leaflet)
 
mapStates = map("state", fill = TRUE, plot = FALSE)
mapStates$value = runif(63,0,1)
pal =  colorNumeric("YlOrRd", domain = mapStates$value)
 
cities <- read.csv(textConnection("
City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994
"))
 
EX1 = leaflet(data = mapStates,780,420) %>%
  addProviderTiles(providers$CartoDB.Positron, group="View 1") %>%
  addProviderTiles("Stamen.TonerLite", group = "View 2") %>%
  addProviderTiles("Thunderforest.Landscape", group = "View 3") %>%
 
  addPolygons(fillColor = ~pal(value), stroke = FALSE, group = "AARP IR") %>%
  addCircles(lng=cities$Long, lat=cities$Lat, weight=1, radius=sqrt(cities$Pop)*30,
             popup = cities$City, group = "City Popul.") %>%
 
  addLayersControl(
    baseGroups = c("View 1", "View 2", "View 3"),
    overlayGroups = c("AARP IR", "City Popul."),
    options = layersControlOptions(collapsed = TRUE)
  ) %>%
  addLegend("bottomleft", pal = pal, values = ~value,
          title = "AARP IR", opacity = .5
  ) %>%
  hideGroup("City Popul.")
 
EX1
```



# Interactive Visualization

## Highcharts Demo

This is an R highcharter library example.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(highcharter)
data(citytemp)
CT <- highchart(width=780, height=400) %>%
  hc_xAxis(categories = citytemp$month) %>%
  hc_add_series(name="Tokyo", data=citytemp$tokyo) %>%
  hc_add_series(name="London", data=citytemp$london) %>%
  hc_add_series(name="Berlin", data=citytemp$berlin) %>%
  hc_add_series(name="NY", data=citytemp$new_york) %>%
  hc_yAxis(title = list(text = "Average Temperature")) %>%
  hc_chart(type="column", options3d=list(enabled=TRUE, beta=15, alpha=15))
CT
```


## Highcharts Demo

This is an R highcharter library example.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
AI <- highchart(width=780, height=400) %>% 
  hc_title(text = "Motor Trend Car Road Tests") %>% 
  hc_add_series_scatter(mtcars$hp, mtcars$mpg, mtcars$wt, mtcars$drat,
                       rownames(mtcars),
                       dataLabels = list(enabled = TRUE, format = "{point.label}")) %>% 
  hc_chart(zoomType = "xy") %>% 
  hc_yAxis(title = list(text = "MPG")) %>%
  hc_xAxis(title = list(text = "Horsepower")) %>%
  hc_tooltip(useHTML = TRUE,
             headerFormat = "<table>",
             pointFormat = paste("<tr><th colspan=\"1\"><b>{point.label}</b></th></tr>",
                                 "<tr><th>Horsepower</th><td>{point.x} hp</td></tr>",
                                 "<tr><th>MPG</th><td>{point.y} mpg</td></tr>",
                                 "<tr><th>Weight(size)</th><td>{point.z} lb/1000</td></tr>",
                                 "<tr><th>Drat(color)</th><td>{point.valuecolor} </td></tr>"),
             footerFormat = "</table>")
AI
```

## Dygraphs Demo

This is an R dygraphs library example.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dygraphs)
library(xts)
data(economics, package = 'ggplot2')
ecm = as.data.frame(economics[,c("date","psavert","uempmed")])
ecm.ts = xts(ecm[,-1], order.by=ecm[,1])

TS <- dygraph(ecm.ts, width=780, height=400,
              main="US personal saving rate and median duration of unemployment") %>% 
  dyRangeSelector() %>% dyLegend(width = 270) %>%
  dyOptions(colors = RColorBrewer::brewer.pal(2, "Set2"))
TS
```



# Rmarkdown Basics

## R Markdown - Paragraphs

This an R Markdown presentation. 
Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. 

For more instructions see <http://rmarkdown.rstudio.com>.

*This sentence is italic.*

**This sentence is bold.**


## R Markdown - Bullets

- Bullet 1  
    \- This could be a paragraph or subtitle.  
    \- This could be a paragraph or subtitle.  
    
- Bullet 2

- Bullet 3


## R Markdown - Smaller Bullets {.smaller}

- Bullet 1  
    \- This could be a paragraph or subtitle.  
    \- This could be a paragraph or subtitle.  
    
- Bullet 2

- Bullet 3


## R Markdown - Notes

We defined the `add` function to compute the sum of two numbers.

This is a code block.
```
This text is displayed verbatim / preformatted
```


## R Markdown - Equations {.smaller}

In this paragraph we have some inline equation $a^2 + b^2$.

This is a LaTex equation.
$$ c^w_i[F(\lambda_r)-1] + c^m_iF(\lambda_i) + \frac {df_i^b(\lambda_i)} {d\lambda_i}			
= (c^w_i + c^m_i) F(\lambda_i) - c^w_i + \frac {df_i^b(\lambda_i)} {d\lambda_i}. $$


