---
title: "Interactive Data Visualization with R"
author: "Chengjun Hou"
date: "March 17, 2015"
output:
  html_document:
    theme: cosmo
---



Today we are going to go through following items:

- RStudio IDE.
- Data structure in R.
- Basic data processing with **dplyr**.
- Static plotting with **ggplot2**.
- Interactive plotting with **rCharts**.
- RMarkdown basics.
- Web application with **shiny**.



## Data Structure
Major data structure in R are *vector*, *data.frame*, and *list*.

The basic data structure in R is the *vector*. There are four common types of atomic vectors: logical, integer, double (often called numeric), and character. Atomic vectors are usually created with `c()`, short for combine:
```{r}
dbl_var = c(1, 2.5, 4.5)
int_var = c(1L, 6L, 10L)
log_var = c(TRUE, FALSE, T, F)
chr_var = c("these are", "some strings")
```

A *list* is a vector but its elements can be of any type, including lists. You construct lists by using `list()`.
```{r}
Lst = list( name="Fred", wife="Mary", no.children=3, child.ages=c(4,7,9) )
Lst
```

A *data.frame* is the most common way of storing data in R, and if used systematically makes data analysis easier. Under the hood, a data frame is a list of equal-length vectors. This makes it a 2-dimensional structure, so it shares properties of both the matrix and the list. This means that a data frame has `names()`, `colnames()`, and `rownames()`, although `names()` and `colnames()` are the same thing. The `length()` of a data frame is the length of the underlying list and so is the same as `ncol()`; `nrow()` gives the number of rows.

You create a data frame using `data.frame()`, which takes named vectors as input:
```{r}
df = data.frame( Name=c("Alice", "Becka", "James", "Jeffrey", "John"),
                 Sex=c("F", "F", "M", "M", "M"),
                 Age=c(13, 13, 12, 13, 12),
                 Height=c(56.5, 65.3, 57.3, 62.5, 59.0),
                 Weight=c(84.0, 98.0, 83.0, 84.0, 99.5),
                 stringsAsFactors = FALSE )
head(df)
```



## Data Processing
The **dplyr** package makes the steps of data manipulation as fast and easy as possible.

To explore the basic data manipulation verbs of **dplyr**, we¡¯ll start with the built in `nycflights13` data frame. This dataset contains all 336776 flights that departed from New York City in 2013. The data comes from the US Bureau of Transportation Statistics, and is documented in `?nycflights13`.
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(nycflights13)
str(flights)
```

`filter()` allows you to select a subset of the rows of a data frame. The first argument is the name of the data frame, and the second and subsequent are filtering expressions evaluated in the context of that data frame. For example, we can select all flights on January 1st with:
```{r}
filter(flights, month == 1, day == 1)
```

`arrange()` works similarly to `filter()` except that instead of filtering or selecting rows, it reorders them. It takes a data frame, and a set of column names (or more complicated expressions) to order by. If you provide more than one column name, each additional column will be used to break ties in the values of preceding columns:
```{r}
arrange(flights, year, month, desc(day))
```

Often you work with large datasets with many columns where only a few are actually of interest to you. `select()` allows you to rapidly zoom in on a useful subset using operations that usually only work on numeric variable positions. And A common use of `select()` is to find out which values a set of variables takes. This is particularly useful in conjunction with the `distinct()` verb which only returns the unique values in a table.
```{r}
distinct(select(flights, origin, dest))
```

As well as selecting from the set of existing columns, it's often useful to add new columns that are functions of existing columns. This is the job of `mutate()`, note that if you only want to keep the new variables, use `transmute()`:
```{r}
transmute(flights,
  gain = arr_delay - dep_delay,
  gain_per_hour = gain / (air_time / 60)
)
```



## Static Plotting
The **ggplot2** package focuses on producing statistical, or data, graphics, but it is unlike most other graphics packages because it has a deep underlying grammar.

In the following example, we split the complete dataset into individual planes and then summarize each plane by counting the number of flights `count = n()` and computing the average distance `dist = mean(Distance, na.rm = TRUE)` and delay `delay = mean(ArrDelay, na.rm = TRUE)`. We then use ggplot2 to display a scatter plot:
```{r}
library(ggplot2)
by_tailnum = group_by(filter(flights, month==1), tailnum)
delay = summarise(by_tailnum,
                   count = n(),
                   dist = mean(distance, na.rm = TRUE),
                   delay = mean(arr_delay, na.rm = TRUE))
delay = filter(delay, count > 20, dist < 2000)
ggplot(delay[-1,], aes(x=dist, y=delay)) +
  geom_point(aes(color = count)) + scale_colour_gradientn(colours=topo.colors(10))
```

After that, we want to compare the arrival delay between American Airlines and Alaska Airlines with a histogram.
```{r}
flights.ft = filter(flights, carrier=="AA"|carrier=="AS")
ggplot(flights.ft, aes(x=arr_delay, fill=carrier)) +
  geom_histogram(aes(y = ..density..), binwidth=1, alpha=.5, position="identity") + xlim(-70,250)
```



## Interactive Plotting
**rCharts** is an R package to create, customize and publish interactive javascript visualizations from R using a familiar lattice style plotting interface. We can install **rCharts** from **github** using the **devtools** package:
```{r, eval=FALSE}
require(devtools)
install_github('rCharts', 'ramnathv')
```
```{r setup, echo=FALSE, results='asis'}
library(rCharts)
add_lib_assets <- function(lib, cdn = F,css=NULL) {
  assets = get_assets(get_lib(lib), cdn = cdn)
  if(!is.null(css)){assets$css=c(assets$css,css)}
  styles <- lapply(assets$css, function(style) {
    sprintf("<link rel='stylesheet' href=%s>", style)
  })

  scripts <- lapply(assets$jshead, function(script) {
    sprintf("<script type='text/javascript' src=%s></script>", script)
  })
  cat(paste(c(styles, scripts), collapse = "\n"))
}
# get assets from online repositories 
add_lib_assets("NVD3",cdn=TRUE,css="http://rawgithub.com/ramnathv/rCharts/master/inst/libraries/nvd3/css/rNVD3.css") 
```

Now we will demonstrate d3js library, NVD3, which produces amazing interactive visualizations with little customization. With each bar indicating the number of people with a specific hair color and eye color, we can choose to group or stack the bar. Also, the chart's legend is interactive, we can click the bubble next to each eye color to turn off and on its bar.
```{r nvd3mbar, results='asis', comment=NA}
hair_eye_female <- filter(as.data.frame(HairEyeColor), Sex == "Female")
n1 <- nPlot(Freq ~ Hair, group = "Eye", data = hair_eye_female, type = "multiBarChart")
n1$chart(color = c('brown', 'blue', '#594c26', 'green'))
n1$print("nvd3mbar")
```

Then we demonstrate the US economic time series from 1967 to 2007. We have `psavert` indicating personal savings rate, and `uempmed` indicating number of unemployed in thousands. We can have a finner display on the top by selecting a range at the bottom. 
```{r nvd3ts, results='asis', comment=NA}
data(economics, package = 'ggplot2')
ecm <- reshape2::melt(economics[,c('date', 'uempmed', 'psavert')], id = 'date')
p7 <- nPlot(value ~ date, group = 'variable', data = ecm, type = 'lineWithFocusChart')
p7$chart(color = c('#FF6666', '#66B2FF'))
p7$xAxis( tickFormat="#!function(d) {return d3.time.format('%b %Y')(new Date( d * 86400000 ));}!#" )
p7$x2Axis( tickFormat="#!function(d) {return d3.time.format('%Y')(new Date( d * 86400000 ));}!#" )
p7$print("nvd3ts")
```



## Resources

- RStudio website <http://www.rstudio.com/>
- Introduction to **dplyr** <http://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html>
- **ggplot2** examples <http://docs.ggplot2.org/0.9.3.1/index.html>
- Introduction to **rCharts** <http://ramnathv.github.io/rCharts/>
- Rmarkdown website <http://rmarkdown.rstudio.com/>
- **shiny** web application gallery <http://shiny.rstudio.com/gallery/>



------
[Interactive Data Visualization with R](https://github.com/chengjunhou/OSU-Informs-Tutorial/)
@ OSU Informs 2015